Unsupervised embedding of trajectories captures the latent structure of mobility

### 2021-02-19 paper/rebuttal/rebuttal.tex

\newcommand{\mycaption}[2]{%
  \caption[#1]{\textbf{#1} \small#2}%
}
% comment out when submitting
\definecolor{myblue}{RGB}{33,150,243}
\definecolor{mygreen}{RGB}{76, 175, 80}
\definecolor{purple}{RGB}{170, 0, 255}

\makeatletter
\renewcommand{\maketitle}{\bgroup\setlength{\parindent}{0pt}
\begin{flushleft}
\Large  \textbf{\@title}
\end{flushleft}\egroup
}
\makeatother
\title{Re: Nature submission 2020-10-19515}
\date{}
\begin{document}
\maketitle

\bigskip
\noindent Dear Dr. Sutherland and Referees, \\

We thank all the referees for their thoughtful and helpful comments and critiques, all of which provided valuable insights.
We acknowledge that our initial submission had a substantial gap regarding the choice of the \textit{word2vec} model, its connection with the gravity model of mobility, and its evaluation. 
While reflecting on these comments and our manuscript, we have made a theoretical breakthrough that addresses the primary issues;
namely, we have discovered a mathematical connection between the \textit{word2vec} (skip-gram) model and the gravity model. 
Furthermore, our subsequent experiments show that our approach's performance is robust, exceeding other baselines. 
Therefore, with these changes, we are confident that our paper is significantly stronger, and its contribution more wide-reaching and impactful.
Thus, we would like to request a reconsideration of our manuscript. 
Again, we are incredibly appreciative of these comments, which provided the insight and motivation needed to make these breakthroughs.

@@

We have developed the mathematical and theoretical basis of our approach for embedding mobility, and made additional updates to our experiments and manuscript, addressing the three main issues raised by referees:

\begin{enumerate}
	\itemsep0em
	\item Our approach and analysis lacked theoretical justification and grounding, and the application of \textit{word2vec} to mobility was arbitrary
	\item Performance of \textit{word2vec} is not well understood in the context of other, simpler embedding techniques
	\item Our findings were not sufficiently grounded in the literature of neural embeddings and mobility
\end{enumerate}

Addressing the first point, we have identified an underlying mathematical equivalence between the Gravity Model of Mobility and the neural embedding technique \textit{word2vec}.
Namely, we demonstrate that \textit{word2vec} 's objective function is equivalent to the gravity law and thus \textit{word2vec} learns gravity-like relationships between entities, whether it be language or locations.
This development constitutes a major breakthrough in our understanding of neural embeddings, and provides theoretical justification for their application to mobility.

For the second point, we have also conducted experiments comparing the performance of neural embedding to an embedding derived by a singular value decomposition of the mobility network, and of a Laplacian eigenmap embedding.
Additionally, we compare \textit{word2vec} model to another matrix factorization approach, which \textit{word2vec} approximates. We find that the embedding distance from \textit{word2vec} outperforms all baselines in explaining and predicting real-world mobility.

### 2021-02-19 paper/rebuttal/rebuttal.tex

Kind Regards,

Dakota, Jisung, Sadamori, Rodrigo, Woo-Sung, Staša, and YY

\clearpage
\tableofcontents
\clearpage

\section{Theoretical justification of \textit{word2vec} and the Gravity Model}
\label{sec:theory}
Why does \textit{word2vec} with the skip-gram with negative sampling model (SGNS) work so well to model human mobility?
Here, we demonstrate that the SGNS model and the gravity model of mobility are mathematically equivalent.
This connection provides strong theoretical justification to our choice of \textit{word2vec}, and constitutes a major breakthrough of our understanding of neural embeddings.
In addition, this motivates the use of SGNS for future study of human mobility.

\subsection{\textit{word2vec} algorithm}
Imagine a trajectory, denoted by ($a_{1}, a_{2}, \ldots, a_{T}$), where $a_{t}$ is the $t$th location in the trajectory. A location, $a_{t}$, is considered to have context locations, $a_{t-w}, \ldots, a_{t-1}, a_{t+1},\ldots, a_{t+w}$, that appear in the window surrounding $a_t$ up to a time lag of $w$, where $w$ is the window size parameter truncated at $t - w \geq 0$ and $t + w \leq T$. 
Then, the SGNS model learns probability $p(a_{t + \tau} \vert a_{t})$, where $-w\leq \tau\leq w$ and $\tau \neq 0$,  by maximizing its log likelihood given by
%
% MAIN W2V Equation
%
\begin{equation}
	\frac{1}{T}\sum_{t = 1}^{T} \sum_{-w \leq \tau \leq w, \tau \neq 0} \log p(a_{t + \tau} \vert a_{t}),
\end{equation}
where,

%
% CONDITIONAL PROBABILITY EQ
%
\begin{equation}
	p(j \given i) = \frac{\exp(\bm{u}_j \cdot \bm{v}_{i})}{Z_i}, \label{eq:cond_prob_w2v}
\end{equation}
where $\bm{v}$ and $\bm{u}$ are the ``in-vector'' and ``out-vector'', respectively,  $Z_i=\sum_{j' \in \mathcal{A}} \exp(\bm{u}_{j'} \cdot \bm{v}_{i})$ is a normalization constant, and $\mathcal{A}$ is the set of all locations. In general, calculating $Z_i$ is computationally expensive and there are two common approximation:
hierarchical softmax \autocite{morin2005hierarchical} and negative sampling \autocite{mikolov2013word2vec}.
Due to its simplicity and performance, negative sampling is the most widely used strategy, which we also adopt in our study.

\subsection{An implicit bias in the negative sampling}
Although negative sampling is the most common approximation, it is known that it is a biased estimator~\autocite{Chia2010,Dyer2014} that optimizes an objective function different from {\it word2vec}.

Negative sampling trains {\it word2vec} using a binary classification task as follows.
A target-context word pair $(i,j)$ is sampled from the given data as a positive pair, and $k$ negative pairs $(i,\ell)$ are sampled from a noise distribution $p_0(\ell)=P^\gamma(\ell)$. ($P(j)$ is the fraction of $j$ in the data, and $\gamma>0$ is a parameter).
Then, given a target-context word pair $(i,j)$, the model predicts whether it is a positive or negative sample using a sigmoid function:
\begin{align}
	\label{eq:sigmoid}
	\frac{1}{1 + \exp\left( - \bm{u}_j \cdot \bm{v}_i \right)}.
\end{align}

This procedure does not guarantee that the embedding optimally converges to the original conditional probability from Eq.~\ref{eq:cond_prob_w2v}, even when increasing the training samples and iterations~\autocite{Chia2010,Dyer2014} because negative sampling has implicit bias.
This problem is addressed by the noise contrastive estimation (NCE) \autocite{Chia2010}, which fits the following probability model,
\begin{align}
	\label{eq:nce}
	f(x) / \sum_{x'} f(x'),
\end{align}
where $f(x)$ is the unnormalized probability for an instance $x$. As in negative sampling, NCE fits the model based on a binary classification task, with positive and negative samples randomly drawn from the given data and a noise distribution $p_0$, respectively, using a bias-corrected sigmoid function:
\begin{align}
	\label{eq:sigmoid2}
	\frac{1}{1 + \exp\left( -f(x) + \log p_0(x) \right)}.
\end{align}

Rewriting the sigmoid function for the negative sampling (Eq.~\ref{eq:sigmoid}) where the bias-corrected sigmoid function yields
\begin{align}
	\label{eq:sigmoid3}
	\frac{1}{1 + \exp\left[ - \tilde f(x) + \log p_0(x)  \right]},
\end{align}
where we have defined
\begin{align}
	\label{eq:unnormalized}
	\tilde f(x) = f(x) p_0(x).
\end{align}
So, the model with the negative sampling optimizes $\tilde f(x)$, not $f(x)$. In case of the {\it word2vec}, the instance is the dot similarity $x=\bm{u}_i \cdot \bm{v}_j$,  $f(x)=\exp(x)$, and $p_0(j)=P^\gamma(j)$, and the SGNS word2vec actually optimize the following conditional probability:
\begin{align}
	P\left(j \given i \right):= \frac{P^\gamma (j)\exp(\bm{u}_j \cdot \bm{v}_{i})}{Z_i},
	\label{eq:real_cond}
\end{align}
where $Z_i=\sum_{j' \in \mathcal{A}} P^\gamma (j') \exp(\bm{u}_{j'} \cdot \bm{v}_{i})$.

@@

\subsection{\textit{word2vec} and the gravity model}

Taking into account the implicit bias in the negative sampling  and setting $w=1$, the \textit{word2vec} predicts the flow between locations with
\begin{align}
	\hat{T}_{ij}= P(i) P(j\given i) = \frac{P(i)P^\gamma (j)\exp(\bm{v}_j \cdot \bm{v}_{i})}{Z_i}. \label{eq:prob_w2v_ng0}
\end{align}
where we have redefined $Z_i=\sum_{j' \in \mathcal{A}} P^ \gamma (j') \exp(\bm{u}_{j'} \cdot \bm{v}_{i})$.

Parameter $\gamma=1$ is a special choice that ensures that, when the embedding dimension is sufficiently large, there exists optimial in-vectors and out-vectors such that $\bm{u}_i = \bm{v}_i$ \autocite{levy2014neural}.
Setting $\gamma = 1$ and substituting $\bm{u}_i = \bm{v}_j$ into Eq.~\ref{eq:prob_w2v_ng0}, the flow predicted by {\it word2vec} is given by
\begin{align}
	\hat{T_{ij}}= P(i) P(j\given i) = \frac{P(i)P(j)\exp(\bm{v}_j \cdot \bm{v}_{i})}{Z_i}. \label{eq:flow_w2v_ng1}
\end{align}

The flow $\hat{T}_{ij}$ is symmetric (\ie $\hat{T}_{ij}=\hat{T}_{ji}$) becausethe skip-gram model neglects whether the context $j$ appears before or after the target $i$ in the trajectory.
If we swap $i$ and $j$ in Eq.~\ref{eq:flow_w2v_ng1}, the numerator remains the same but the denominator can be different.
Therefore, to ensure $\hat{T}_{ij} = \hat{T}_{ji}$, the denominator $Z_i$ should be a constant.

Taken together, the {\it word2vec} model with the negative sampling predicts a flow in the same form as in the gravity model:
\begin{align}
	\hat{T_{ij}}=  C P(i) P(j)  \exp(\bm{v}_j \cdot \bm{v}_{i}), \label{eq:flow_w2v_ng0}
\end{align}
where $C$ is a positive constant.

In other words, \emph{word2vec with the negative sampling is a equivalent to a gravity model}, with the mass given by the location's frequency $P(i)$, and the distance measured by their dot similarities.
While the gravity model predicts mobility flows from the given mass and locations, {\it word2vec} finds the mass and locations that best explain the given mobility flow.
Therefore, the embedding distance is inherently tied to mobility flow, and hence, has greater predictive power than the geographic distance.

In practice, because we do not have unlimited amounts of data (which is also noisy) and the optimization may not find the true optimum, the idealized result may not hold. 
Indeed, we tend to find that the in- and out-vectors tend to be different and that the cosine similarity tends to better capture real-world mobility than the dot similarity.
Nevertheless, a model with dot similarity is still the second best after cosine similarity (Tables~\ref{supp:table:r2_table},~\ref{supp:table:rmse_exp_table},~\ref{supp:table:rmse_power_table}), and the embedding distance still outperforms all alternatives we considered.

\subsection{Changes to the manuscript}
To reflect our new findings, we have made the following changes to the manuscript:
\begin{itemize}
	\itemsep0em
	\item Following results demonstrating the performance of the embedding distance compared to geographic distance, we add the section \textbf{word2vec and the gravity model}, which summarizes the link between the two models.
	\item We also added the section \textbf{An implicit bias in the negative sampling} to the methods section of our manuscript.
	\item Since the formulation of the gravity used in the above derivations assumes $\gamma = 1$, we use $\gamma = 1$ for all reported results. No results have been substantively changed.
	\item We added a sentence to the paper speculating on why cosine similarity works best.
	\item We also added a sentence noting the in-vector and out-vector issue, and speculating on reasons/consequences.
\end{itemize}

### 2021-02-19 paper/rebuttal/rebuttal.tex

%
% SECTION-PERFORMANCE COMPARISONS
%
\section{Performance of \textit{word2vec} compared with baselines}
\subsection{Alternatives considered}
In our original manuscript, we compared the performance of what we call the embedding distance---the cosine distance between vectors learned by \textit{word2vec}---to a range of alternatives.
These alternatives included the geographic distance between locations, the distance most commonly used in definitions of the gravity model of mobility, along with the dot product similarity between embedding vectors, and the cosine distance and Jensen-Shannon divergence between vectors obtained from a personalized-page-ranking procedure on the underlying mobility network.

Following the advice of the referees, we also add additional baselines to illustrate how the embedding distance performs compared to simpler embedding alternatives.
The first of these baselines uses the distance between vectors in an embedding space produced through a singular-value decomposition of the underlying location co-occurrence matrix.
The second baseline is the distance between vectors obtained through a Laplacian eigenmap embedding of the location of the matrix.

In addition to these suggested baselines, we also consider distances derived from embeddings learned through matrix factorization following the procedure outlined by Levy \& Goldberg~\autocite{levy2014neural}.
The \textit{word2vec} model  implicitly approximates this matrix-factorization procedure, and so comparison to these ``Levy'' distances can demonstrate whether the approximation of neural embeddings outperform direct-calculations.

\subsection{New experimental parameters}
In addition to the new baselines, we have also incorporated additional experimental parameters into our analysis.
The first is $\gamma$, a hyper-parameter of the \textit{word2vec} model, which defines the exponent which shapes the negative sampling distribution.
The default of `\texttt{gensim's} implementation of \textit{word2vec} is to set $\gamma = 0.75$, and this was the value used in our original manuscript.
However, to connect \textit{word2vec} to the gravity model of mobility, we assume $\gamma = 1.0$ (see section 1).
We now report all main results using $\gamma = 1.0$, however we also compare its performance to embeddings learned using $\gamma = 0.75$, as we do with the number of embedded dimensions and window size.

@@

We also consider an experimental parameter differentiating between definitions of \textit{population}, in the context of scientific mobility.
Our original manuscript defined the population of an organization as its average annual number of unique mobile and non-mobile authors who publish with that organizational affiliation.
However, because non-mobile authors do not have mobility trajectories, and are thus not included in training of the embedding, this definition of population introduces additional, potentially confounding information into the experiment.
To tease apart the impact of population definitions, we consider two additional definitions.
The first defines an organization's population as the average annual number of unique mobile authors, only, who publish with the organization.
The second definition defines the population akin to language, as simply its frequency across all mobility trajectories.

\subsection{Result}

%
% Figure---Metric performance
%
\begin{figure}[ht!]
	\centering
	\includegraphics[width=0.95\textwidth]{\figdir/Descriptive/distance_metric_performance.pdf}
	\mycaption
	{Neural embeddings outperform baselines for scientific mobility.}
	{
		Cosine distance between embedding vectors generated with \textit{word2vec} explains more of the flux, and better predicts flux when used in the gravity model of mobility than geographic distance and other baselines.
		Shown is the correlation between the flux and embedding distances, measured with $R^{2}$ (left), and the prediction error when using the distance as input to the gravity model of mobility.
		For prediction error, we show results based on both the exponential and power-law forms of the gravity model.
		All embedding-based methods use dimensions of 300.
		Embedding distance is obtained from neural embeddings learned with window size of 1 and $\gamma = 1$.
		The metric with the highest correlation and lowest prediction error are marked with an asterisk. 
		In all cases, organization population is defined as the mean annualized number of unique mobile and non-mobile authors, and flux is calculated for all global mobility.
		Baselines include the top-performing distance metrics calculated between vectors obtained by personalized-page rank (PPR), singular value decomposition (SVD), laplacian eigenmap, and direct-factorization~\autocite{levy2014neural} approaches, as well as the geographic distance between organizations.
		Embedding distance better explains and predicts flux than any other baseline, though there is some variation by experimental parameters (Table~\ref{supp:table:r2_table}, Table~\ref{supp:table:rmse_exp_table}, and Table~\ref{supp:table:rmse_power_table}).
	}
	\label{fig:supp:distancemetrics}
\end{figure}

We find that, using the best performing dimensions (300) and window size (1),  $\gamma = 1.0$, and for default experimental parameters (global mobility, population defined using all mobile and non-mobile scholars), that the embedding distance, or cosine distance between vectors learned through neural embedding, explains more of the flux, and leads to better predictions than other baselines (Fig.~\ref{fig:supp:distancemetrics}).
This performance advantage holds when considering both the exponential and the power-law forms of the gravity model of mobility.
Although the euclidean distance between directly-factorized vectors performs comparably, and even narrowly beats out the embedding distance under certain experimental conditions (Tables~\ref{supp:table:r2_table},~\ref{supp:table:rmse_exp_table},~\ref{supp:table:rmse_power_table}); \textit{word2vec} shows the msot consistent and robust performance, as suggested by Levy \& Goldberg~\autocite{levy2014neural}.

\subsection{Why do neural embeddings outperform direct matrix factorization?}
Why do neural-embedding approaches, which rely on stochastic gradient descent, outperform direct matrix factorization, especially given that \textit{word2vec} \emph{is} implicitly approximating factorization?
An issue in the matrix factorization approach is that it is extremely sensitive to small flows between affiliations. The matrix factorization approach embeds affiliations $i$ and $j$ such that their dot similarity is as close as to $\log (T_{ij} / P(i) P(j))$. If the flow $T_{ij}$ is considerably small or zero, the dot similarity goes $-\infty$, pushing $i$ and $j$ very far from the other affiliations in the embedding space~\autocite{levy2014neural, Qui2018}. This is particularly problematic when the window size is small because most affiliation pairs would have no flow, which is indeed the case in our experiments. To circumvent this problem, the previous studies~\autocite{levy2014neural, Qui2018} add a constant flow between the affiliation pairs with no flow. However, this changes the data to which we fit the model, and small flows still have a considerably strong impact on the embedding.

### 2021-02-19 paper/rebuttal/rebuttal.tex

	\begin{itemize}
		\item Liu, Xin, Yong Liu, and Xiaoli Li. ``Exploring the Context of Locations for Personalized Location Recommendations.'' IJCAI. 2016.
		\item Feng, Shanshan, et al. ``POI2Vec: geographical latent representation for predicting future visitors.'' AAAI. 2017.
		\item Yao, Zijun, et al. ``Representing Urban Functions through Zone Embedding with Human Mobility Patterns.'' IJCAI. 2018.
		\item Cao, Hancheng, et al. ``Habit2vec: Trajectory semantic embedding for living pattern recognition in population.'' IEEE Transactions on Mobile Computing 19.5 (2019): 1096-1108.
	\end{itemize}

	However, the novelty of this work is quite narrow, as there are several prior works that develop trajectory-based embeddings of human mobility based on the word2vec algorithm. For example, the following references all propose word2vec-inspired embeddings of human mobility data:

	Notably, the location embedding methodology in Liu et al (2016) (see Section 3.1) is essentially the same as the approach used in the submitted manuscript, and the work by Liu et al (2016) is well-known with over 100 citations.
}

\response{Thank you for pointing out the missing references!
	We have carefully read through each, and learned a lot in the process.
	They have now been incorporated into our manuscript's introduction.
	Although there are existing work on trajectory-based mobility embedding, we argue that our new result establishing the mathematical equivalence of \textit{word2vec} and the gravity law represents a novel and substantial contribution. 
}

@@

\rcomment{
	I found the specific use of the SGNS algorithm unmotivated. There are many ways to generate useful low-dimensional embeddings, so why is the SGNS method prioritized above others? This method is popular, but popularity alone is not sufficient justification and I see no theoretical reason why it should be the most appropriate method. For example, a key point in this manuscript is the fact that the the SGNS embeddings capture the flux between locations better than geographic distance. However, if this is such an important point, then why not simply learn a low-dimensional embedding approximation of the flux metric directly? such as SVD?
}

\response{
	Thank you for the suggestion!
	Since our original submission, we have uncovered a mathematical connection between the skip-gram model with negative sampling, and the gravity model of mobility.
	Namely, we find that \emph{SGNS is a gravity model}.
	This finding demonstrates why SGNS performs so well, and demonstrates the utility of the model for studying contexts where gravity relationships are expected, such as in human mobility.
	We refer you to Section 1 of this letter, above, for an investigation of this connection, and how we adjusted our manuscript given this breakthrough.

	Additionally, we compared the performance of SGNS to several baselines, including the distances between vectors obtained with singular-value decomposition, Laplacian eigenmap, and also Levy
	\& Goldberg's~\autocite{levy2014neural} matrix factorization, the last of which directly calculates what SGNS only approximates.
	We find that in nearly every case, SGNS outperforms all of these baselines, further illustrating the potential of such models for embedding and studying mobility.
	We refer you to Section 2 of this rebuttal letter, above, for more detailed results and the changes we made to our manuscript.
}

@@

\rcomment{%
	The methodology has potential limitations in terms of reproducibility and robustness, which should be addressed. This is no specific fault of the authors, but both the SGNS/word2vec algorithm and the UMAP algorithm are inherently stochastic. This means that these approaches involve inherent randomness, and for researchers to replicate the exact effect seen in this manuscript, one would need to ensure the exact same random seeds etc. were used. In my view, this is a serious limitation, especially given that the UMAP results and learned embeddings are reified and treated as first-class objects of scientific study.
}

\response{Thank you for this comment. While robustness is a general concern for these methods, we found that their application to our datasets were in fact quite robust, and observed no evidence that findings were the result of methodological artefacts. This is confirmed in extra work done in response to these reviews.

	We re-ran all of our results using different starting seeds and found all results were qualitatively similar, and almost quantitatively identical.
	In addition, we expanded the set of experimental parameters, training 18 different embeddings with different dimensions, window sizes, and now, $\gamma$.
	We found that while performance varies across each combination of parameters, the main results remain qualitatively similar.
	While we can run more combinations of parameters, and more iterations of \textit{word2vec}, we see no evidence that results will drastically vary.

	We believe that the ratio of the number of organizations represented in our dataset to the total number of mobility trajectories makes our results especially robust.
	About 8,661 organizations are represented across over three million career trajectories of scientists, such that nearly every organization appears enough times, and in enough contexts, to produce a robust vector representation.
	However, this may not hold true for all problem domains—for example, natural language includes many millions of tokens, many of which will occur only a few times, allowing stochasticity to have a much stronger effect on the final representation.
	Heeding this as a general concern, we have added the following sentence to our limitations paragraph in our conclusion:

	“We however, note that all results shown here are robust to this stochasticity, likely emerging from the limited ``vocabulary'' of scientific mobility, airports, and accommodations (several thousand) and the massive datasets used (several million trajectories). Applications of word2vec to problem domains where the ratio of the vocabulary to data is smaller, however, should take caution to ensure that findings are not the result of random fluctuations.”

@@

}

\rcomment{%
	Another interesting point in this work is that the authors chose to not report p-values or the results of statistical significance tests (except confidence intervals). I presume this was a conscious decision, which I do not necessarily agree with. I see no critical error in the approach used by the authors, but I would have found aspects of the manuscript easier to understand in places with results of statistical tests (e.g., associated with goodness-of-fit tests formally test that the embedding-based distances are more correlated with the flux values). Given the sample sizes involved, I have no doubt that the results are statistically significant. However, I personally do see value in rigorously performed frequentist hypothesis testing, and I felt that this manuscript could be improved by employing such techniques.

	Lastly, the authors do not actually state how the 99 confidence intervals were constructed. (E.g., was this via a bootstrap approach?) This detail should be added.
}

\response{Thank you for this feedback.
	You are correct that it was a conscious choice to not report statistical significance in the form of a P-value.
	The use of statistical significance is contentious, and Nature itself has hosted some of this debate~\autocite{amrhein2019statistics}.
	As such, we avoided p-values in our research when it made sense to do so.
	In our analysis, all significance levels were below 0.0001, with most computationally indistinguishable from zero.
	However, we recognize that p-values can help readers interpret results.
	As such, we have opted to add to a sentence such as \say{\textit{Correlation is calculated on the data in the log-log scale ($p < 0.0001$ across all fits)}.} or similar to relevant figures in the main text and supporting materials.

@@

	We have also provided detail about how confidence intervals were calculated in relevant figure captions.
	For example, we state in figure 2's caption that \say{\textit{99\% confidence intervals are plotted for the mean flux in each bin based on a normal distribution.}}, which should give readers enough information to reproduce the confidence intervals following their standard definition.
	Other confidence intervals, such as intervals for the loess regression line in figure 6, are calculated using the smoothing functions in the \texttt{ggplot2} package of R, and refer readers to the relevant documentation.
}

\rcomment{%
	F. Suggested improvements:
	References and discussion of prior work using word2vec-style embeddings of mobility data.
	Additional experiments using alternative embedding methods, especially an SVD-based embedding method, as well as robustness/reproducibility studies across different runs of the SGNS would be essential to strengthen the paper.
	In addition, some additional results demonstrating how the embeddings provide novel scientific insight, rather than just recapitulating known structures, would be essential to make this work above-the-bar for a high-impact publication such as Nature.

}

\response{We believe that our new mathematical results substantially improve the manuscript and extend its contribution and demonstrates novel scientific insight, and justifies the use of word2vec towards mobility data. By adding this, and addressing your other points (above), we believe that the revised manuscript sufficiently remedies your concerns.

### 2021-02-19 paper/current/main.tex
yy comments: section 3.1 rebuttal

In spite of its promise, our study has several limitations.
First, the skip-gram \textit{word2vec} model does not leverage directionality, meaning that embedding will be less effective at capturing mobility for which directionality is critical.
Future studies may also consider bi-directional embeddings, such as BERT~\autocite{devlin2018bert}, to incorporate directionality.
Second, the neural embedding approach is most useful in cases of mobility between discrete geographic units such as between countries, cities, and businesses; this approach is less useful in the case of mobility between locations represented using geographic coordinates, such as in the modeling of animal movements.
Neither of these methodological limitations is insurmountable, and future work can aim to incorporate directionality and identify meaningful representations of continuous mobility data.
Third, neural embeddings are an inherently stochastic procedure, and so results may change across different iterations.
We however, note that all results shown here are robust to this stochasticity, likely emerging from the limited ``vocabulary'' of scientific mobility, airports, and accommodations (several thousand) and the massive datasets used (several million trajectories).
Applications of \textit{word2vec }to problem domains where the ratio of the vocabulary to data is smaller, however, should take caution to ensure that findings are not the result of random artifacts.
Finally, the case of scientific mobility presents domain-specific limitations.
Reliance on bibliometric metadata means that we capture only long-term mobility such as migration, rather than the array or more frequent short-term mobility such as conference travel and temporary visits.
The kinds of mobility we do capture---migration, and co-affiliation---although conceptually different, are treated identically by our model.
Also, our data might further suffer from bias based on publication rates: researchers at prestigious organizations tend to have more publications, leading to these organizations appearing more frequently in affiliation trajectories.

>>>>

In spite of its promise, our study has several limitations.
First, the skip-gram \textit{word2vec} model does not leverage directionality, meaning that embedding will be less effective at capturing mobility for which directionality is critical.
Future studies may also consider bi-directional embeddings, such as BERT~\autocite{devlin2018bert}, to incorporate directionality.
Second, the neural embedding approach is most useful in cases of mobility between discrete geographic units such as between countries, cities, and businesses; this approach is less useful in the case of mobility between locations represented using geographic coordinates, such as in the modeling of animal movements.
Neither of these methodological limitations is insurmountable, and future work can aim to incorporate directionality and identify meaningful representations of continuous mobility data.
Third, neural embeddings are an inherently stochastic procedure, and so results may change across different iterations.
We however, note that all results shown here are robust to this stochasticity, likely emerging from the limited ``vocabulary'' of scientific mobility, airports, and accommodations (several thousand) and the massive datasets used (several million trajectories).
Applications of \textit{word2vec }to problem domains where the ratio of the vocabulary to data is smaller, however, should take caution to ensure that findings are not the result of random fluctuations.
Finally, the case of scientific mobility presents domain-specific limitations.
Reliance on bibliometric metadata means that we capture only long-term mobility such as migration, rather than the array or more frequent short-term mobility such as conference travel and temporary visits.
The kinds of mobility we do capture---migration, and co-affiliation---although conceptually different, are treated identically by our model.
Also, our data might further suffer from bias based on publication rates: researchers at prestigious organizations tend to have more publications, leading to these organizations appearing more frequently in affiliation trajectories.
